Multi-domain, multilingual translation with latent multi-task group dropout

### 2022-01-14 2022/acl/acl_latex.tex
local changes

\subsection{System design and configuration}
\subsubsection{Multidomain translation systems}
Our systems for the multidomain experiments are designed as follows:
\begin{itemize}
\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 2048; the multi-head attention mechanism contains 8 heads; 6 layers in the encoder; 6 layers in the decoder.\fyDone{How many layers}
\item Adapter-based Transformer: The intermediate feedforward dimension is set to 2048;% as in \citet{Pham21revisiting}.
\item Transformer using Latent multi-task group dropout (\system{LaMGD} Transformers):\fyDone{Explain acronym} There is no change in the architecture. We group the 512 nodes in each layer into 16 groups of 32 consecutive nodes. For each domain, only 12 out of the 16 groups are selected. The number of parameters of the variational distribution is $L\times k \times L \times n_d$, which is negligible in comparison to the size of the Transformer model.
\item \item Transformer using heuristic multi-task group dropout (\system{HMGD} Transformer): we share 320 nodes for every task, and reserve 32 nodes for each task (making $320 + 32*6 = 512$ nodes in total).
\end{itemize}

@@

The temperature parameter $\tau$ for the $\operatorname{Soft-Top-K}$ operator is gradually decreased from $0.5$ to $0.2$ according to the following policy:
\begin{align*}
\tau = \operatorname{min}\{ 0.2, 0.5 * \exp^{-r*step} \},
\end{align*}
in which $r=0.0001$.
While \citet{Gong21pay,Gong21adaptive} fixed $\tau$ to be 0.2, we select an anneal policy for $\tau$ proposed by previous studies \citep{Jang17categorical}. Finally, we fix the weight of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ to $0.0001$ in the training loss in every experiments

@@

\subsection{Multilingual translation}
Results for the multilingual experiments are in Table~\ref{tab:multilingual}.  The \system{LaMGD} Transformer achieves an improvement of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in the O2M-related, M2O-related, M2O-diverse conditions, respectively. Significant gains are observed for languages \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction) which are very low-resource languages in our sets. However, \system{LaMGD} Transformer is outperformed by the multilingual \system{Transformer} and language Adapters for the O2M-diverse condition.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{9}{r|}} \hline
    O2M-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[91.6m]} & 4.8 &7.3&20.8&21.1&39.7&19.8&22.6&15.2&18.9 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]} &4.3&6.8&21.1&22&39.7&20&23&15.2&19 \\ 
    \system{LaMGD} Transformer  \hfill{\footnotesize[+0m]}  & 5.2&\SB{9.4}&20.6&\SB{22.8}&39.6&19.6&22.4&15.0&19.33 \\ 
	\hline
    \hline
    M2O-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[67.8m]} &11.4&16.6&28.5&	27.1&43.7&24.6&30.3&25.6&25.98 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]} &10.1&15.8&28.4&26.8&43.7&24.5&30.2&25.6&25.64\\ 
    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  &11.3&\SB{17.4}&28.6&\SB{28.7}&43.7&24.5&30.7&25.6&26.31 \\ 
    \hline
    \hline
    O2M-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.2&4&12.7&22.2&31.8&\SB{34.0}&38.3&8.3&20.19 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]}  &10.2&4&\SB{13.3}&21.9&32.2&\SB{34.1}&38.5&8.3&20.31 \\ 
    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}&10.1&3.8&12.6&\SB{22.8}&31.8&33.4&38.1&8.1&20.09\\
    \hline 
    \hline
    M2O-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.4&9.7&20.5&31.8&37.5&38.7&39.8&19.0&27.43 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]}  &22.5&9.4&20.0&30.6&37.2&38.2&39.3&19.0&27.03\\ 
    \system{LaMGD} Transformer  \hfill{\footnotesize[+0m]} &\SB{23.5}&9.6&\SB{21.5}&32.2&37.7&38.6&40.0&18.9&27.75 \\
    \hline
  \end{tabular}
  \caption{Experiments with multilingual translation. Boldface denotes significant gains over \system{Transformer}.\fyDone{What is the bold for ?}}
  \label{tab:multilingual}
\end{table*}

>>>>

\subsection{System design and configuration}
\subsubsection{Multidomain translation systems}
Our systems for the multidomain experiments are designed as follows:
\begin{itemize}
\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 2048; the multi-head attention mechanism contains 8 heads; 6 layers in the encoder; 6 layers in the decoder.\fyDone{How many layers}
\item Adapter-based Transformer: The intermediate feedforward dimension is set to 2048;% as in \citet{Pham21revisiting}.
\item Transformer using Latent multi-task group dropout (\system{LaMGD} Transformers):\fyDone{Explain acronym} There is no change in the architecture. We group the 512 nodes in each layer into 16 groups of 32 consecutive nodes. For each domain, only 12 out of the 16 groups are selected. The number of parameters of the variational distribution is $L\times k \times L \times n_d$, which is negligible in comparison to the size of the Transformer model.
\item Transformer using heuristic multi-task group dropout (\system{HMGD} Transformer): we share 320 nodes for every task, and reserve 32 nodes for each task (making $320 + 32*6 = 512$ nodes in total).
\end{itemize}

@@

The temperature parameter $\tau$ for the $\operatorname{Soft-Top-K}$ operator is gradually decreased from $0.5$ to $0.2$ according to the following policy:
\begin{align*}
\tau = \operatorname{min}\{ 0.2, 0.5 * \exp^{-r*step} \},
\end{align*}
in which $r=0.0001$.
While \citet{Gong21pay,Gong21adaptive} fixed $\tau$ to be 0.2, we combine it with an anneal policy for $\tau$ proposed by previous studies \citep{Jang17categorical} to encourage the diversity in the beginning of the training. Finally, we fix the weight of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ to $0.0001$ in the training loss in every experiments.

@@

\subsection{Multilingual translation}
Results for the multilingual experiments are in Table~\ref{tab:multilingual}.  The \system{LaMGD} Transformer achieves an improvement of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in the O2M-related, M2O-related, M2O-diverse conditions, respectively. Significant gains are observed for languages \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction) which are very low-resource languages in our sets. However, \system{LaMGD} Transformer is outperformed by the multilingual \system{Transformer} and language Adapters for the O2M-diverse condition.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{9}{r|}} \hline
    O2M-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[91.6m]} & 4.8 &7.3&20.8&21.1&39.7&19.8&22.6&15.2&18.9 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]} &4.3&6.8&21.1&22&39.7&20&23&15.2&19 \\ 
    \system{LaMGD} Transformer  \hfill{\footnotesize[+0m]}  & 5.2&\SB{9.4}&20.6&\SB{22.8}&39.6&19.6&22.4&15.0&19.33 \\ 
	\hline
    \hline
    M2O-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[67.8m]} &11.4&16.6&28.5&	27.1&43.7&24.6&30.3&25.6&25.98 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]} &10.1&15.8&28.4&26.8&43.7&24.5&30.2&25.6&25.64\\ 
    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  &11.3&\SB{17.4}&28.6&\SB{28.7}&43.7&24.5&30.7&25.6&26.31 \\ 
    \hline
    \hline
    O2M-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.2&4&12.7&22.2&31.8&\SB{34.0}&38.3&8.3&20.19 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]}  &10.2&4&\SB{13.3}&21.9&32.2&\SB{34.1}&38.5&8.3&20.31 \\ 
    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}&10.1&3.8&12.6&\SB{22.8}&31.8&33.4&38.1&8.1&20.09\\
    \hline 
    \hline
    M2O-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.4&9.7&20.5&31.8&37.5&38.7&39.8&19.0&27.43 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]}  &22.5&9.4&20.0&30.6&37.2&38.2&39.3&19.0&27.03\\ 
    \system{LaMGD} Transformer  \hfill{\footnotesize[+0m]} &\SB{23.5}&9.6&\SB{21.5}&32.2&37.7&38.6&40.0&18.9&27.75 \\
    \hline
  \end{tabular}
  \caption{Experiments with multilingual translation. Boldface denotes significant gains over \system{Transformer} ($p=0.05$).\fyDone{What is the bold for ?}}
  \label{tab:multilingual}
\end{table*}
